Toward single particle reconstruction without particle picking: Breaking the detection limit

### 2018-06-06 TEX/draft4.tex
Furhter minor changes

We here explicitly derive algebraic relations between the autocorrelation functions of the micrographs and the autocorrelation functions of the target signals. For ease of exposition, we do so in the 1-D case. Extension to the 2-D case is straightforward. We also give additional technical details regarding the two experiments presented in Section~\ref{sec:results}.

@@

In general, for a signal $z$ of length $m$, the autocorrelation of order $q \geq 1$ is given for any integer shifts $\ell_1, \ldots, \ell_{q-1}$ by
\begin{align}
	a_z^q[\ell_1,\ldots,\ell_{k-1}]  & = \frac{1}{m} \sum_{i=-\infty}^{+\infty} z[i]z[i+\ell_1]\cdots z[i+\ell_{q-1}],
	\label{eq:ac_general}
\end{align}
where indexing of $z$ out of the bounds $0, \ldots, m-1$ is zero-padded.
For our purposes, this will be applied both to $x_k$'s (each of length $L$) and to $y$ (of length $N$).
Explicitly, the first-, second- and third-order autocorrelations are given by
\begin{align} 
	a_z^1 & = \frac{1}{m} \sum_{i=0}^{m-1} z[i], \nonumber\\
	a_z^2[\ell] & = \frac{1}{m} \sum_{i = \max\{0, -\ell\}}^{m-1 + \min\{0, -\ell\}} z[i]z[i+\ell], \nonumber\\
	a_z^3[\ell_1,\ell_2] & = \frac{1}{m} \sum_{i = \max\{0, -\ell_1, -\ell_2\}}^{m-1 + \min\{0, -\ell_1, -\ell_2\}} z[i]z[i+\ell_1]z[i+\ell_2]. \label{eq:ac_special}
\end{align}
The autocorrelation functions have symmetries. Specifically, $a_z^2[\ell] = a_z^2[-\ell]$, and
\begin{align*}
	a_z^3[\ell_1,\ell_2] = a_z^3[\ell_2,\ell_1]=a_z^3[-\ell_1,\ell_2-\ell_1].
\end{align*}
%Additionally, if the moments of the signal depend only on the difference between the indices (Toeplitz structure), then they are equivalent to the autocorrelation functions.
Taking these symmetries into consideration, one can show that $a_z^1, a_z^2$ and $a_z^3$ contain \TODO{TKTK} (respectively) non-trivial distinct real numbers, generically.

@@

For the special case $K = 1$ where a single signal signal $x = x_1$ must be recovered, the relation between autocorrelations of the micrograph and those of $x$ is particularly simple, so that we treat it first. It is useful to introduce some notation: let $M$ denote the number of occurrences of $x$ in $y$, and let
\begin{align}
	\gamma & = \frac{M L}{N}
\end{align}
denote the density of $x$ in $y$ (that is, the fraction of entries of $y$ occupied by occurrences of $x$.) The spacing constraint~\eqref{eq:spacing} imposes $\gamma\leq\frac{L}{2L-1}\approx 1/2$.

@@

Returning to the general case $K \geq 1$, let $M_1, \ldots, M_k$ denote the number of occurrences of signals $x_1, \ldots, x_K$ respectively, and define
\begin{align}
	\gamma_k & = \frac{M_k L}{N}, & \gamma & = \sum_{k=1}^K\gamma_k.
\end{align}
As above, we consider the asymptotic regime where $M_1,\ldots,M_K,N\to\infty$ while preserving the ratios $\gamma_k$ constant.
% 
Still under the spacing constraint~\eqref{eq:spacing}, similarly to the developments above, one can estimate a mixture of the autocorrelations of the $K$ target signals from the autocorrelations of the micrograph:
\begin{align}
\lim_{N\to\infty} a_y^1 & = \sum_{k=1}^K\gamma_k a_{x_k}^1, \nonumber\\
\lim_{N\to\infty} a_y^2[\ell] & = \sum_{k=1}^K\gamma_k a_{x_k}^2[\ell] +\sigma^2\delta[\ell],  \label{eq:data_ac}\\
\lim_{N\to\infty} a_y^3[\ell_1,\ell_2] & = \sum_{k=1}^K\gamma_k a_{x_k}^3[\ell_1,\ell_2] + \sigma^2\left(\sum_{k=1}^K\gamma_k a_{x_k}^1\right)(\delta[\ell_1,0]+\delta[0,\ell_2]+\delta[\ell_1,\ell_2]), \nonumber
\end{align}
where $\ell, \ell_1, \ell_2 = 0, \ldots, L-1$. The left hand side is straightforward to estimate from data. The right hand side involves polynomial functions of unknowns $\gamma_1, \ldots, \gamma_K, x_1, \ldots, x_K$. The task is to solve these polynomial equations in a robust way.

>>>>

We derive algebraic relations between the autocorrelation functions of the micrographs and the autocorrelation functions of the target signals. For ease of exposition, we do so in the 1-D case. Extension to the 2-D case is straightforward. We also give additional technical details regarding the two experiments presented in Section~\ref{sec:results}.

@@

In general, for a signal $z$ of length $m$, the autocorrelation of order $q \geq 1$ is given for any integer shifts $\ell_1, \ldots, \ell_{q-1}$ by
\begin{align}
	a_z^q[\ell_1,\ldots,\ell_{k-1}]  & = \frac{1}{m} \sum_{i=-\infty}^{+\infty} z[i]z[i+\ell_1]\cdots z[i+\ell_{q-1}],
	\label{eq:ac_general}
\end{align}
where indexing of $z$ out of the range $0, \ldots, m-1$ is zero-padded.
For our purposes, this will be applied both to $x_k$'s (each of length $L$) and to $y$ (of length $N$).
Explicitly, the first-, second- and third-order autocorrelations are given by
\begin{align} 
	a_z^1 & = \frac{1}{m} \sum_{i=0}^{m-1} z[i], \nonumber\\
	a_z^2[\ell] & = \frac{1}{m} \sum_{i = \max\{0, -\ell\}}^{m-1 + \min\{0, -\ell\}} z[i]z[i+\ell], \nonumber\\
	a_z^3[\ell_1,\ell_2] & = \frac{1}{m} \sum_{i = \max\{0, -\ell_1, -\ell_2\}}^{m-1 + \min\{0, -\ell_1, -\ell_2\}} z[i]z[i+\ell_1]z[i+\ell_2]. \label{eq:ac_special}
\end{align}
The autocorrelation functions have symmetries. Specifically, $a_z^2[\ell] = a_z^2[-\ell]$, and
\begin{align*}
	a_z^3[\ell_1,\ell_2] = a_z^3[\ell_2,\ell_1]=a_z^3[-\ell_1,\ell_2-\ell_1].
\end{align*}
%Additionally, if the moments of the signal depend only on the difference between the indices (Toeplitz structure), then they are equivalent to the autocorrelation functions.
Taking these symmetries into consideration, one can show that $a_z^1, a_z^2$ and $a_z^3$ contain \TODO{TKTK} (respectively) non-trivial distinct real numbers, generically.

@@

For the special case $K = 1$ where a single signal $x = x_1$ must be recovered, the relation between autocorrelations of the micrograph and those of $x$ is particularly simple, so that we treat it first. It is useful to introduce some notation: let $M$ denote the number of occurrences of $x$ in $y$, and let
\begin{align}
	\gamma & = \frac{M L}{N}
\end{align}
denote the density of $x$ in $y$ (that is, the fraction of entries of $y$ occupied by occurrences of $x$.) The spacing constraint~\eqref{eq:spacing} imposes $\gamma\leq\frac{L}{2L-1}\approx 1/2$.

@@

Returning to the general case $K \geq 1$, let $M_1, \ldots, M_K$ denote the number of occurrences of signals $x_1, \ldots, x_K$ respectively, and define
\begin{align}
	\gamma_k & = \frac{M_k L}{N}, & \gamma & = \sum_{k=1}^K\gamma_k.
\end{align}
As above, we consider the asymptotic regime where $M_1,\ldots,M_K,N\to\infty$ while preserving the ratios $\gamma_k$ constant.
% 
Still under the spacing constraint~\eqref{eq:spacing}, similarly to the developments above, one can estimate a mixture of the autocorrelations of the $K$ target signals from the autocorrelations of the micrograph:
\begin{align}
\lim_{N\to\infty} a_y^1 & = \sum_{k=1}^K\gamma_k a_{x_k}^1, \nonumber\\
\lim_{N\to\infty} a_y^2[\ell] & = \sum_{k=1}^K\gamma_k a_{x_k}^2[\ell] +\sigma^2\delta[\ell],  \label{eq:data_ac}\\
\lim_{N\to\infty} a_y^3[\ell_1,\ell_2] & = \sum_{k=1}^K\gamma_k a_{x_k}^3[\ell_1,\ell_2] + \sigma^2\left(\sum_{k=1}^K\gamma_k a_{x_k}^1\right)(\delta[\ell_1,0]+\delta[0,\ell_2]+\delta[\ell_1,\ell_2]), \nonumber
\end{align}
where $\ell, \ell_1, \ell_2 = 0, \ldots, L-1$. The left hand side is straightforward to estimate from data: it provides a succinct summary of it. The right hand side involves polynomial functions of unknowns $\gamma_1, \ldots, \gamma_K, x_1, \ldots, x_K$. The task is to solve these polynomial equations in a robust way.
